Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2021-08-17 article/draft.tex
updates in seection 1 for FB and TW

Mainstream social media platforms may suspend the account of a specific user when they deem that the platforms' rules have been violated. Account suspension can be temporary or permanent.  When the suspension is temporary the user is prohibited for a limited period of time from posting content on their account, but created content prior to suspension remains available to the user and their followers. However, when the suspension is permanent, in most cases, followers or subscribers have no longer access to the content prior to the suspension and the user can no longer use the account to create new content.  In what follows, we focus on the implementation of this policy by several platforms and provide simple examples to illustrate.

@@

When an account is permanently suspended by Facebook, it disappears from the platform. That is,  the data can no longer be scrapped and it also disappears from the CrowdTangle API.\footnote{CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages.} Facebook publishes on monthly basis a {\it coordinated inauthentic behavior} report, where it informs how many personal accounts, pages or groups were deleted and to which {\it deceptive network} they may have belonged to.\footnote{See the \href{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}{April 2021 report} as an example.} But as long as external persons do not have access to deleted accounts data, these reports cannot be verified by independent researchers or journalists.

@@

Facebook can also apply a temporary suspension, and in this case the data can often be collected and analyzed. For example, \href{https://www.facebook.com/DonaldTrump/}{Donald Trump’s official Facebook page}  has been suspended following the Capitol attack on January 6, 2021.\footnote{See \href{https://www.facebook.com/zuck/posts/10112681480907401}{https://www.facebook.com/zuck/posts/10112681480907401}} Nevertheless the page’s data is still present in the CrowdTangle API. Thus, after manually adding this page to the CrowdTangle dashboard, we collected the $6$ $083$ posts it had published between January $1$, $2020$ and June $15$, $2021$ using the {\it posts} endpoint.\footnote{See the endpoint documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Posts}{https://github.com/CrowdTangle/API/wiki/Posts}.} We used Minet command line tool \cite{minet} to collect the data. We can verify on figure \ref{fig1_fb} that the {\it Donald J. Trump} page has not published any content since January $6$, 2021, and that this behavior is not consistent with the page’s previous activity: an average of $16$ posts were published each day on Facebook before the suspension.

@@

%paragraph strike system + say following example is a manual decision. 
Twitter has implemented a strike system as part of their Civic Integrity Policy and their COVID-19 misleading information policy. Violations of both policies can entail strikes, where two strikes lead to a 12-hour account lock and five or more strikes lead to permanent suspension from the platform. A list of notable Twitter temporary and permanent suspensions can be found on  \href{https://en.wikipedia.org/wiki/Twitter_suspensions}{Wikipedia}. The 12-hour account lock is hard to observe in the data, especially for users who do not have an over the clock tweeting activity. In this section, we provide one example of a temporary suspension of a Twitter account, that seems to be the result of a manual decision concerning a Tweet which violated the rules.

The Twitter account $@LifeSite$ of the website lifesitenews.com has been suspended for at least two periods of time: from end of 2019 until fall 2020 for 308 days, then again since January 2021 for having violated Twitter Rules\footnote{See Lifesitenews's article discussing the reason for the suspension: \href{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}. Twitter rules can be found at: \href{https://help.twitter.com/en/rules-and-policies/twitter-rules}{https://help.twitter.com/en/rules-and-policies/twitter-rules}. }. In particular, this website has several failed fact-checks concerning the published articles, according to Iffy.news.\footnote{See \href{https://mediabiasfactcheck.com/life-site-news/}{https://mediabiasfactcheck.com/life-site-news/}}. We collected the activity (tweets, replies, quotes, retweets) on their Twitter account via the Twitter API, using the historical search endpoint. We then plotted the number of Tweets, Retweets, Quotes and Replies per day, as shown in panel $a$ of figure \ref{fig2}). The two periods of temporary suspension are clearly observed in the data as the user(s) of the account were not allowed to use the functionalities of the Twitter Platform.

@@

To further assess the impact of this double temporary suspension, we collect via Minet Command Line Tool\cite{minet}, all the tweets that have shared during the same period a url link containing lifesitenews.com. Panel $(b)$ of figure \ref{fig2}, shows that during both periods of temporary suspension, other users still shared lifesitenews.com links and that the level was only slightly below the tweeting and retweeting levels prior to the first temporary suspension. More specifically, there was an average of 960 tweets (including retweets) per day over the first temporary suspension period of 308 days from December 9, 2019 until October 12, 2020, against an average of 977 tweets (including retweets) per day during the exact same period one year earlier. Finally, panel $(b)$ points towards the limitations of suspending an account to limit the spread of its content.

>>>>

Mainstream social media platforms may suspend the account of a specific user when they deem that the platforms' rules have been violated. Account suspension can be temporary or permanent.  When the suspension is temporary the user is prohibited for a limited period of time from posting content on their account, but created content prior to suspension remains available to the user and their followers. However, when the suspension is permanent, in most cases, followers or subscribers have no longer access to the content prior to the suspension and the user can no longer use the account to create new content.  In what follows, we focus on the implementation of this policy by Facebook, Twitter and YouTube, and we provide simple examples to illustrate.

@@

When an account is permanently suspended by Facebook, it disappears from the platform. That is,  the data can no longer be scrapped and it also disappears from the CrowdTangle API.\footnote{CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages.} Facebook publishes on monthly basis a {\it coordinated inauthentic behavior} report, where it informs how many personal accounts, pages or groups were deleted and to which {\it deceptive network} they may have belonged to.\footnote{See the \href{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}{April 2021 report} as an example.} 
But as long as external persons do not have access to historical data of deleted accounts, these reports cannot be verified by the academic community, NGOs and journalists.

@@

Facebook can also apply a temporary suspension, and in this case the data can often be collected and analyzed. For example, \href{https://www.facebook.com/DonaldTrump/}{Donald Trump’s official Facebook page}  has been suspended following the Capitol attack on January 6, 2021.\footnote{See \href{https://www.facebook.com/zuck/posts/10112681480907401}{https://www.facebook.com/zuck/posts/10112681480907401}} Nevertheless the page’s data is still present in the CrowdTangle API. 
%Thus, after manually adding this page to the CrowdTangle dashboard, 
Hence, we collected the $6$ $083$ posts it had published between January $1$, $2020$ and June $15$, $2021$ using the {\it posts} endpoint.\footnote{See the endpoint documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Posts}{https://github.com/CrowdTangle/API/wiki/Posts}.} We used Minet command line tool \cite{minet} to collect the data. We can verify on figure \ref{fig1_fb} that the {\it Donald J. Trump} page has not published any content since January $6$, 2021, and that this behavior is not consistent with the page’s previous activity: an average of $16$ posts were published each day on Facebook before the suspension.

@@

%paragraph strike system + say following example is a manual decision. 
Twitter has implemented a strike system as part of their Civic Integrity Policy and their COVID-19 misleading information policy. Violations of both policies can entail strikes, where two strikes lead to a 12-hour account lock and five or more strikes lead to permanent suspension from the platform. A list of notable Twitter temporary and permanent suspensions can be found on  \href{https://en.wikipedia.org/wiki/Twitter_suspensions}{Wikipedia}.\footnote{See \href{https://en.wikipedia.org/wiki/Twitter\_suspensions}{https://en.wikipedia.org/wiki/Twitter\_suspensions}.}
The 12-hour account lock is hard to observe in the data, 
especially for users who do not have an over the clock regular tweeting activity.\footnote{The following tool \href{https://makeadverbsgreatagain.org/allegedly/}{https://makeadverbsgreatagain.org/allegedly/} provides an overview of the daily and hourly tweeting activity, including repetition of Tweets for any given Twitter account.} 
In this section, we provide one example of a temporary suspension of a Twitter account, that seems to be the result of a manual decision concerning a Tweet which violated the rules.

The Twitter account $@LifeSite$ of the website lifesitenews.com has been suspended for at least two periods of time: from end of 2019 until fall 2020 for 308 days, then again since January 2021 for having violated Twitter Rules\footnote{See Lifesitenews's article discussing the reason for the suspension: \href{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}. Twitter rules can be found at: \href{https://help.twitter.com/en/rules-and-policies/twitter-rules}{https://help.twitter.com/en/rules-and-policies/twitter-rules}. }. In particular, this website has several failed fact-checks concerning the published articles, according to Iffy.news and feedback.org.\footnote{See \href{https://mediabiasfactcheck.com/life-site-news/}{https://mediabiasfactcheck.com/life-site-news/} and \href{https://open.feedback.org/media/S4}{https://open.feedback.org/media/S4}.}. We collected the activity (tweets, replies, quotes, retweets) on their Twitter account via the Twitter API, using the historical search endpoint. We then plotted the number of Tweets, Retweets, Quotes and Replies per day, as shown in panel $a$ of figure \ref{fig2}). The two periods of temporary suspension are clearly observed in the data as the user(s) of the account were not allowed to use the functionalities of the Twitter Platform.

@@

To further assess the impact of this double temporary suspension, we collect via the Twitter API v2, all the tweets that have shared during the same period a url link containing lifesitenews.com. Panel $(b)$ of figure \ref{fig2}, shows that during both periods of temporary suspension, other users still shared lifesitenews.com links and that the level was only slightly below the tweeting and retweeting levels prior to the first temporary suspension. More specifically, there was an average of 960 tweets (including retweets) per day over the first temporary suspension period of 308 days from December 9, 2019 until October 12, 2020, against an average of 977 tweets (including retweets) per day during the exact same period one year earlier.

Finally, suspending a Twitter account is an intervention which aims at penalizing users, in order to make them respect the Twitter rules. In the case of Twitter accounts linked to domain names, the activity of the account usually consists in spreading on social media published articles on the website. Panel $(b)$ points towards the limitations of suspending the Twitter account $@LifeSite$, because the content created by the domain lifesitenews.com was still being actively shared by other Twitter users.
